---
title: "probability_distribution_fitting"
output: html_document
date: "2026-06-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(ISLR)
data(College)
```


We consider the variable `Accept`, which gives the number of applications accepted by each college.

```{r}
accept <- College$Accept
summary(accept)
```

## Part 1: Histogram of accepted applications with exponential fit

We assume that the number of applications accepted can be modelled as an exponential random variable with parameter beta equal to the sample mean.

```{r}
beta_hat <- mean(accept)
beta_hat
```

The exponential density in R uses `rate = 1 / beta`.

```{r}
hist(accept,
     probability = TRUE,
     breaks = 30,
     main = "Histogram of Accepted Applications with Exponential Fit",
     xlab = "Number of Applications Accepted")

x_vals <- seq(min(accept), max(accept), length.out = 1000)
y_vals <- dexp(x_vals, rate = 1 / beta_hat)

lines(x_vals, y_vals, lwd = 2)
```

## Part 1: Histogram of log accepted applications with normal fit

Now consider the log of the number of applications accepted.

```{r}
log_accept <- log(accept)

mu_hat <- mean(log_accept)
sigma_hat <- sd(log_accept)

mu_hat
sigma_hat
```

We now fit a normal distribution to the log data.

```{r}
hist(log_accept,
     probability = TRUE,
     breaks = 30,
     main = "Histogram of Log Accepted Applications with Normal Fit",
     xlab = "Log Number of Applications Accepted")

x_vals <- seq(min(log_accept), max(log_accept), length.out = 1000)
y_vals <- dnorm(x_vals, mean = mu_hat, sd = sigma_hat)

lines(x_vals, y_vals, lwd = 2)
```

## Answer

The histogram of the original `Accept` variable is strongly right-skewed, with most colleges accepting relatively small numbers of applications and a few colleges accepting very large numbers. The exponential density captures the general decreasing shape, but it does not fit the full distribution perfectly, especially because of the long right tail.

After taking logs, the histogram of `log(Accept)` is much more symmetric and approximately bell-shaped. The fitted normal density follows the shape of the log-transformed data reasonably well. This suggests that the normal model for `log(Accept)` provides a better fit than the exponential model for the original `Accept` values.


## Part 2: Probability that a college accepts more than 2500 applications

Under the fitted exponential model:

```{r}
p_exp <- 1 - pexp(2500, rate = 1 / beta_hat)
p_exp
```

Under the fitted normal model for the log data:

```{r}
p_norm_log <- 1 - pnorm(log(2500), mean = mu_hat, sd = sigma_hat)
p_norm_log
```

Observed proportion in the data:

```{r}
p_observed <- mean(accept > 2500)
p_observed
```

Compare the model predictions with the observed data:

```{r}
p_exp
p_norm_log
p_observed
```

```{r}
abs(p_exp - p_observed)
abs(p_norm_log - p_observed)
```
## Answer
The fitted exponential model gives probability 0.2899.
The fitted normal model on the log data gives probability 0.2353.
The observed proportion in the data is 0.2432.

To compare the accuracy of the two fitted models, we calculate the absolute difference between each model prediction and the observed proportion:

|0.2899 - 0.2432| = 0.0466

|0.2353 - 0.2432| = 0.0080

Since (0.0080 < 0.0466), the fitted normal model on the log data is more accurate. 
The exponential model overestimates the probability that a college accepts more than 2500 applications, while the normal model on the log data gives a prediction much closer to the observed proportion.
